---
title: "Peer Assessment 1 for Reproducible Research"
author: "Emre"
date: "2017 M10 8"
output: 
  html_document: 
    fig_caption: yes
    keep_md: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

This is to downlaod "data.zip"and extract "activity.csv" if not available  
Read data from file  
Convert string format to date  

```{r prepocess}
csvFile <- "activity.csv"
if( !file.exists(csvFile) ){
  zipFile <- "data.zip"
  # Download if not exists
  if( !file.exists(zipFile) ){
    download.file("https://github.com/emrsen/RepData_PeerAssessment1/blob/master/activity.zip", destfile = zipFile)
  }
  # Extract zip
  unzip(zipFile, exdir = ".") 
}

# Read data from file
df <- read.csv(csvFile, stringsAsFactors = F)
# Covert string to date format
df$date2 <- as.Date(df$date,"%Y-%m-%d")
```

## What is mean total number of steps taken per day?

Calculating daily sum of the steps

```{r mean_total_steps_daily}
# Calculate daily sum of steps
stepsday<-tapply(df$steps, df$date2, FUN=sum, na.rm=TRUE)
#Plot histogram for the daily sum of the steps
barplot(stepsday,main = "Total steps per day distribution",xlab = "Daily steps")
abline(h=mean(stepsday))
abline(h=median(stepsday), col="red")
hist(stepsday,main = "Total steps per day histogram distribution",xlab = "Daily steps")
#Calculate mean of daily sum of steps black line
mean(stepsday)
#Calculate median of daily sum of steps red line
median(stepsday)
```

## What is the average daily activity pattern?

Graphing average daily activity pattern interval basis

```{r graph_average}
#Calculate average steps for each inteval
stepsmeanint <- tapply(df$steps, df$interval, FUN = mean, na.rm = TRUE)
#Plot graph for the average steps
plot(names(stepsmeanint),stepsmeanint, type = "l", main = "Steps in Time Series Plot", xlab = "5-Minute Intervals", ylab = "Average Number of Steps")
#maximum average steps for the time of the day 5 minute interval
as.numeric(names(which.max(stepsmeanint)))
#maximum average steps values
max(stepsmeanint)

```

## Imputing missing values

Replacing missing values using calculated average mean of each steps

```{r impute_missing_values}
#create subset of missing steps
dfna <- df[which(is.na(df$steps)),]
#create a new copy of the data frame df
df_imputed <- df
# Replace missing steps in the new imputed data frame with the calculated missing steps
df_imputed$steps[is.na(df$steps)] <- stepsmeanint[as.character(dfna$interval)]
# Calculate daily sum of imputed steps
stepsday_imp<-tapply(df_imputed$steps, df_imputed$date2, FUN=sum, na.rm=TRUE)
# Plot histogram for the daily sum of the steps
barplot(stepsday_imp,main = "Total steps(imputed) per day distribution",xlab = "Daily steps")
abline(h=mean(stepsday_imp))
abline(h=median(stepsday_imp), col="red")
hist(stepsday_imp,main = "Total steps per day (imputed) histogram distribution",xlab = "Daily steps")
# Calculate mean of daily sum of imputed steps black line
mean(stepsday_imp)
# Calculate median of daily sum of imputed steps red line
median(stepsday_imp)
# Mean and median value increased and equalised. 
# After imputing the missing values with average daily values
```

## Are there differences in activity patterns between weekdays and weekends?

Displaying weekdays and weekends patterns

```{r weekdayend_compare}
#Adding weekday info to the imputed data
df_imputed$WD <- weekdays(df_imputed$date2)
#Adding Weekday or weekend info
df_imputed$WDF <- ifelse(df_imputed$WD %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"), "WeekDay", "WeekEnd")
#Comparale graphics in ggplot
library(ggplot2)
ggplot(df_imputed, aes(x=interval, y=steps, color=WDF)) + geom_line()

```
  
Weekdays people walk more early morning hours  
Weekends people walk more in afternoon hours